Real-Time Scheduling Algorithms and Battery Consumption of Mobile Devices

### 2017-05-08 paper-RMDON.tex
primer draft

In this paper, we present [a set of] [the following] real-time scheduling algorithms  [A, B, C] that have been proposed to improve energy-saving for mobile devices. We discuss their main features

\section{Hard Real-Time versus Soft Real-Time Systems}
The following are the major differences between hard and soft real-time systems according to \cite{juvva}. 
\paragraph{Response time requirements} on hard real-time systems are in the order of milliseconds or less and can result in a catastrophe if not met. In contrast, the response time requirements of soft real-time systems are higher and not very strict.

@@

\section{Real-Time Task Scheduling Algorithms}
Several schemes of classification of real-time task scheduling algorithms exist. A scheme in \cite{mall} classifies the real-time task scheduling algorithms based on how the scheduling points (the points on time line at which the scheduler makes decisions regarding which task is to be run next) are defined. According to this classification scheme, the three main types of schedulers are: clock-driven, event-driven, and hybrid. The clock-driven schedulers are those in which the scheduling points are determined by the interrupts received from a clock. In the event-driven ones, the scheduling points are defined by certain events which precludes clock interrupts. The hybrid ones use both clock interrupts as well as event occurrences to define their scheduling points.

@@

%Several variations to these two basic algorithms exist.

%We  focus  on  a  preemptive hard real-time systems in which periodic real-time tasks are scheduled with the Earliest-Deadline-First (EDF) algorithm or the Rate-Monotonic (RM) algorithm, the two most widely used real-time system models\cite{w-s-liu}.

@@

\section{DVS Algorithms}
There are two kinds of voltage scheduling approaches for hard real-time systems depdending on the voltage scaling granularity: intra-task DVS (IntraDVS) and inter-task DVS (InterDVS). The intra-task DVS algorithms adjust the voltage within an individual task boundary, while the inter-task DVS algorithms determine the voltage on a task-by-task basis at each scheduling point. The main difference between them is whether the slack times are used for the current task or for the tasks that follow.  InterDVS algorithms distribute the slack times from the current task for the following tasks, while IntraDVS algorithms use the slack times from the current task for the current task itself.

\subsection{Intra-task DVS algorithms} 
In  scheduling hard real-time tasks, in order to guarantee the timing constraint of each task,  the execution times of tasks are usually assumed to be the worst case execution times (WCETs).  However, since a task has many possible execution paths, there are large execution time variations among them. So, when the execution path taken at run time is not the worst case execution path (WCEP), the task may complete its execution before its WCET, resulting in a slack time. In that case, IntraDVS exploits such slack times and adjusts the processor speed. IntraDVS algorithms can be classified into two types depending on how to estimate slack times and how to adjust speeds.

\subsubsection{Path-based method} the voltage and clock speed are determined based on a predicted reference execution path, such as WCEP. For example, when the actual execution deviates from the predicted reference execution path, the clock speed is adjusted.  If the new path takes significantly longer to complete its execution than the reference path, the clock speed is raised to meet the deadline constraint. On the other hand, if the new path can finish its execution earlier than the reference path, the clock speed is lowered to reduce the energy consumption. Program locations for possible speed scaling are identified using static program analysis [16] or execution time profiling [11].

\subsubsection{Stochastic method}
based on the idea that it is better to start the execution at a low speed and accelerate the execution later when needed than to start with a high speed and reduce the speed later when slack time is found. By starting at a low speed, if the task finishes earlier than its WCET, it does not need to execute at a high speed. Theoretically, if the probability density function of execution times of a task is known \emph{a priori}, the optimal speed schedule can be computed [5].  Under the stochastic method, the clock speed is raised at specific time instances, regardless of the execution paths taken. The stochastic IntraDVS may not utilize all the potential slack times.

\subsection{Inter-task DVS algorithms}
InterDVS algorithms exploit the ``run-calculate-assign-run'' strategy to determine the supply voltage, which can be summarized as follows: (1) run a current task, (2) when the task is completed, calculate the maximum allowable execution time for the next task, (3) assign the supply voltage for the next task, and (4) run the next task.  Most InterDVS algorithms differ during step (2) in computing the maximum allowed time for the next task $\tau$ which is the sum of WCET of $\tau$ and the slack time available for $\tau$.

A generic InterDVS algorithm consists of two parts: slack estimation and slack distribution.  The goal of the slack estimation part is to identify as much slack times as possible while the goal of the slack distribution part is to distribute the resulting slack times so that the resulting speed schedule is as uniform as possible. Slack times generally come from two sources; \emph{static slack times} are the extra times available for the next task that can be identified statically, while {dynamic slack times} are caused from run-time variations of the task executions.

\subsubsection{Slack estimation methods}
\emph{Static slack estimation Maximum constant speed:} One of the most commonly used static slack estimation methods is to compute the maximum constant speed, which is defined as the lowest possible clock speed that guarantees the feasible schedule of a task set [18].  For example, in EDF scheduling, if the worst case  processor utilization (WCPU) $U$ of a given task set is lower than 1.0 under the maximum speed $f_{max}$, the task set can be scheduled with a new maximum speed $f'_{max} = U \cdot f_{max}$. Although more complicated, the maximum constant speed can be statically calculated as well for RM scheduling [18, 5].

@@

\emph{Stretching to NTA:} even though a given task set  is scheduled with the maximum constant speed, since the actual execution times of tasks are usually much less than their WCETs, the tasks usually have dynamic slack times.  One simple method to estimate the dynamic slack time is to use the  arrival  time  of  the  next  task   [18]. (The  arrival  time of the next task is denoted by NTA.) Assume that the current task $\tau$ is scheduled at time $t$. If NTA of $\tau$ is later than $(t + \textnormal{WCET}(\tau))$, task $\tau$ can be executed at a lower speed so that its execution completes exactly at the NTA.

%Figure   1   shows   examples   of   the Stretching-to-NTA method.  When a single task $\tau$ is activated as shown in Figure 1(a), the execution of $\tau$ can be stretched to NTA. When multiple tasks are activated, there can be several alternatives in stretching options.  For example, the dynamic slack time may be given to a single task or distributed equally to all activated tasks.  Cases I and II of Figure 1(b) illustrate these two options, respectively.

\emph{Priority-based slack stealing}: this method exploits the basic  properties of  priority-driven scheduling such as  RM and EDF. The basic idea is that when a higher-priority task completes its execution earlier than its WCET, the following lower-priority tasks  can use the  slack time from the completed higher-priority task.  It is also possible for a higher-priority task to utilize the slack times from completed lower-priority tasks.  However, the latter type of slack stealing is computationally expensive to implement precisely.  Therefore, the existing algorithms are based on heuristics [2, 10].

\emph{Utilization updating:} the actual processor utilization during run time is usually lower than the worst case processor utilization. The utilization updating technique estimates the required processor performance at the current scheduling point by recalculating the expected worst case processor utilization using the actual execution times of completed task instances [14].   When the processor utilization is updated, the clock speed can be adjusted accordingly. The main merit of this method is its simple implementation, since only the processor utilization of completed task instances have to be updated at each scheduling point.

\paragraph{Slack distribution methods} In distributing slack times, most InterDVS algorithms have adopted a greedy approach, where all the slack times are given to the next activated task. This  approach is not an optimal solution, but the greedy approach is widely used because of its simplicity.

\section{Representative Algorithms} \label{hard-rt-algs}
Many DVS algorithms have been proposed or developed, especially for hard real-time systems. Since lowering the supply voltage also decreases the maximum achievable clock speed, various DVS algorithms for hard real-time systems have the goal of reducing supply voltage dynamically to the lowest possible level while satisfying the tasks' timing constraints\cite{kim-2}. The following is a representative list of DVS algorithms that have been proposed for addressing energy efficiency on mobile devices: two of which are based on the RM scheduling policy, while the other algorithms are based on the EDF scheduling policy, the two most widely used real-time system models\cite{w-s-liu}. Two algorithms were selected for IntraDVS, one from path-based IntraDVS algorithms, and the other from stochastic methods.

@@

\textbf{==> TENTATIVO: hablar de lpfpsEDF que viene en Power Concious...}

Shin's intra-task DVS algorithm \cite{shin-kim-lee} and Gruian's algorithm \cite{gruian} are used as representative IntraDVS algorithms which use path-based method and the stochastic method, respectively.

%Pillai \emph{et. al.} also proposed two other DVS algorithms [14], \texttt{ccEDF} and \texttt{laEDF}, for EDF scheduling policy. These algorithms estimate slack time of a task using the utilization updating method. While \texttt{ccEDF} adjusts the voltage and clock speed based on run-time variation in processor utilization alone, laEDF takes a more aggressive approach by estimating the amount of work required to be completed before NTA.

%\texttt{DRA} and \texttt{AGR} , which were proposed by Aydin \emph{et. al.} in [2],  are two representative DVS algorithms that are based on the priority-based slack stealing method. The \texttt{DRA} algorithm estimates the slack time of a task using the priority-based slack stealing method along with the maximum constant speed and the Stretching-to-NTA methods.  Aydin \emph{et. al.} also extended the \texttt{DRA} algorithm and proposed another DVS algorithm called \texttt{AGR} for more aggressive slack estimation and voltage/clock scaling. In AGR, in addition to the priority-based slack stealing, more slack times are identified by computing the amount of work required to be completed before NTA. \texttt{lpSEH} is another  DVS  algorithm  which  is  based  on the priority-based slack stealing method [10]. Unlike \texttt{DRA} and \texttt{AGR}, \texttt{lpSEH} extends the priority-based slack stealing method by adding a procedure that estimates the slack time from lower-priority tasks that were completed earlier than expected. \texttt{DRA}, \texttt{AGR}, and \texttt{lpSEH} algorithms are somewhat similar to one another in the sense that all of them use the maximum constant speed in the off-line phase and the Stretching-to-NTA method in the on-line phase in addition to the priority-based slack stealing method. For IntraDVS algorithms, Shin's intra-task DVS  algorithm [16] (intraShin) and Gruian's' algorithm [5] (intraGruian) are used as representative algorithms of the path-based method and the stochastic method, respectively.
%
%\texttt{darEDF} is based on an efficient slack task scheduling scheme that employs dynamic speed setting slack in run queue. The algorithm uses a dynamic version of the average rate heuristic first and delays slack absorption, resulting in better battery performance.

\section{Results}
In \cite{kim-2} a performance comparison of most of the algorithms presented in section \ref{hard-rt-algs} was made using a simulation environment named SimDVS. This work is relevant because it is argued that although each DVS algorithm is shown to be quite effective in reducing the energy/power consumption of a target system under its own experimental scenarios, these DVS algorithms have not been quantitatively evaluated under a unified framework, making it a difficult task for low-power embedded system developers to select an appropriate DVS algorithm for a given application/system.

%Although each DVS algorithm is shown to be quite effective in reducing the energy/power consumption of a target system under its own experimental scenarios,  these recent DVS algorithms have not been quantitatively evaluated under a unified framework, making it a difficult task for low-power embedded system developers to select an appropriate DVS  algorithm for  a  given application/system. A quantitative analysis of the energy-efficiency is particularly important because most of these DVS algorithms are based on both static and dynamic slack analysis techniques whose performance is difficult to predict analytically.  In addition, their energy  efficiency  fluctuate  significantly  depending  on  the workload variations, task set characterizations, and execution paths taken, further requiring a quantitative comparison study.

%It is argued that the energy efficiency of InterDVS algorithms depends significantly of slack estimation and the appropriateness of slack distribution. The following highlights the results obtained when evaluating InterDVS algorithms.

\paragraph{Impact of the number of task on the energy consumption} as the number of tasks increases,  the  energy  efficiency  of \texttt{lppsEDF}, \texttt{lppsRM}, and \texttt{ccRM} that only use the Stretching-to-NTA technique do not significantly improve, while that of the other more aggressive InterDVS algorithms improves significantly when compared with the energy consumption of an applicacion running on a DVS-unaware system with a power-down mode only. In the Stretching-to-NTA method, the slack time that can be exploited is limited to the time between the completion of a task instance and the arrival time of the next task instance, which is largely independent of the number of tasks in the system. For the other InterDVS algorithms, since the slack times can be taken from any completed task instance, as the number of task increases, each task has more slack sources and can be scheduled with a lowered clock speed.

\paragraph{Worst case processor utilization of task set} except for \texttt{lppsEDF}, the energy consumption of InterDVS algorithms increases as a linear function of WCPU of a task  set. For \texttt{lppsEDF}, the energy consumption increases faster than a linear function of WCPU of a task set. This indirectly indicates that the dynamic slack estimation method of \texttt{lppsEDF} is not very effective. 
\texttt{lppsEDF} shows better energy efficiency than \texttt{ccEDF} when WCPU is less than 0.7. In \texttt{ccEDF}, the clock speed is determined using the actual processor utilization at the scheduling point.  Since the actual processor utilization increases when a low-speed task instance completes its execution, the next task instance needs to be executed in a higher speed. Such  voltage fluctuation occurs more often as the WCPU decreases. Thus, as the WCPU decreases, the energy efficiency of ccEDF becomes worse than that of \texttt{lppsEDF}. The results for \texttt{lppsRM} and \texttt{ccRM} are very similar to that of \texttt{lppsEDF}.

\paragraph{Speed bound} the energy efficiency of \texttt{AGR} and \texttt{lpSEH} is very close to the theoretical lower bound\footnote{Determined by the Yao's algorithm in \cite{yao}} when the speed bound factor is near 0.5. For the aggressive InterDVS algorithms, the energy efficiency is highest when the speed bound factor was set to average case processor utilization (ACPU). When the selected speed bound factor is close to ACPU $(= 0,55 \times \textnormal{WCPU})$, the best energy effciency is achieved for \texttt{laEDF}. For \texttt{ccEDF} this trend does not hold. For RM InterDVS algorithms, the performance gap between the energy efficiency and that of the theoretical lower bound was roughly 35$\sim$40\%.

@@

\subsection{Performance evaluation of Intra-Task DVS algorithms}
The two representative IntraDVS algorithms perform quite differently depending on available slack times. When the relative energy consumption ratio of intraGruian over intraShin was measured, if the ratio is larger than 1, intraGruian performs better than intraShin. When  the  slack  ratio  is  less  than  1.2, intraShin outperforms intraGruian because intraShin spends more time in the lower speed region than intraGruian. When the slack ratio is increased, intraGruian spends more time in the lower speed region than intraShin.

\section{Related Work}
Although this work is focused in hard real-time scheduling algorithms, CPU scheduling and energy/power efficiency of soft real-time systems and mobile devices has been addressed as well. In \cite{yuan-nahrstedt-1} a framework to integrate DVS into soft real-time scheduling for open mobile systems is presented. Its goal is to achieve energy saving DVS while preserinf resource guarantees of soft real-time scheduling. In \cite{yuan-nahrstedt} an enery efficient soft real-time CPU scheduler intended for mobile multimedia systems named \emph{GRACE-OS} is presented.

In \cite{kim-2} hybrid methods are discussed as well. By using this methods researchers try to answer quest of whether hybrid DVS algorithms will perform better than pure InterDVS or pure IntraDVS. The study indicates that the performance of a HybridDVS algorithm can be better than a pure IntraDVS algorithm or a pure InterDVS algorithm. However, the differences in energy efficiency depend on the characteristics of both the IntraDVS and the InterDVS components used in the HybridDVS algorithm.

@@

%Since in these systems resource demands differs from the ones presented in hard real-time systems, performance improvements are approached depending of the application quality of service (QoS) that is being intended to be improved, for instance disks access, multimedia or network utilization. 
%
%\begin{itemize}
%    \item otros algoritmos para hard-time (lfpsEDF por ejemplo)
%    \item scheduling en mobile devices basado en offloading
%\end{itemize}

\section{Conclusions}
In this paper, real-time scheduling principles and algorithms and their impact on mobile device energy/power consumption are discussed.

@@

>>>>

In this paper, we introduce a set or real-time scheduling algorithms for hard real-time systems have been proposed to improve energy-saving for mobile devices. The rest of the paper is organized as follows. The paper begins pointing the differences between hard real-time and soft real-time systems in Section \ref{sec:hardrt-vs-softrt}. Then, we provide an introduction to real-time task scheduling algorithms in Section \ref{sec:rt-algs}. We focus our attention in \emph{earliest deadline first} and \emph{rate monotonic} algorithms because they have been proven to be the most popular algorithms for hard real-time systems\cite{w-s-liu}. In Section \ref{sec:dvs-algs} the basics of DVS algorithms are presented. Section \ref{sec:rt-algs} presents several variations of hard real-time and DVS algorithms that have been proposed that helps to improve battery performace. Results on Section \ref{sec:results} measure how these algorithms behave under certain conditions. The paper is concluded in Section \ref{sec:conclusion}.

\section{Hard Real-Time versus Soft Real-Time Systems} \label{sec:hardrt-vs-softrt}
The following are the major differences between hard and soft real-time systems according to \cite{juvva}. 
\paragraph{Response time requirements} on hard real-time systems are in the order of milliseconds or less and can result in a catastrophe if not met. In contrast, the response time requirements of soft real-time systems are higher and not very strict.

@@

\section{Real-Time Task Scheduling Algorithms} \label{sec:rt-algs}
Several schemes of classification of real-time task scheduling algorithms exist. A scheme in \cite{mall} classifies the real-time task scheduling algorithms based on how the scheduling points (the points on time line at which the scheduler makes decisions regarding which task is to be run next) are defined. According to this classification scheme, the three main types of schedulers are: clock-driven, event-driven, and hybrid. The clock-driven schedulers are those in which the scheduling points are determined by the interrupts received from a clock. In the event-driven ones, the scheduling points are defined by certain events which precludes clock interrupts. The hybrid ones use both clock interrupts as well as event occurrences to define their scheduling points.

@@

@@

\section{DVS Algorithms} \label{sec:dvs-algs}
There are two kinds of voltage scheduling approaches for hard real-time systems depdending on the voltage scaling granularity: intra-task DVS and inter-task DVS. The intra-task DVS algorithms adjust the voltage within an individual task boundary, while the inter-task DVS algorithms determine the voltage on a task-by-task basis at each scheduling point. The main difference between them is whether the slack times are used for the current task or for the tasks that follow. Inter-task DVS algorithms distribute the slack times from the current task for the following tasks, while intra-task DVS algorithms use the slack times from the current task for the current task itself.

\subsection{Intra-task DVS algorithms} 
In  scheduling hard real-time tasks, in order to guarantee the timing constraint of each task,  the execution times of tasks are usually assumed to be the worst case execution times (WCETs).  However, since a task has many possible execution paths, there are large execution time variations among them. So, when the execution path taken at run time is not the worst case execution path (WCEP), the task may complete its execution before its WCET, resulting in a slack time. In that case, intra-task DVS exploits such slack times and adjusts the processor speed. Intra-task DVS algorithms can be classified into two types depending on how to estimate slack times and how to adjust speeds.

\subsubsection{Path-based method} the voltage and clock speed are determined based on a predicted reference execution path, such as WCEP. For example, when the actual execution deviates from the predicted reference execution path, the clock speed is adjusted.  If the new path takes significantly longer to complete its execution than the reference path, the clock speed is raised to meet the deadline constraint. On the other hand, if the new path can finish its execution earlier than the reference path, the clock speed is lowered to reduce the energy consumption. Program locations for possible speed scaling are identified using static program analysis\cite{shin-kim-lee} or execution time profiling\cite{lee-sakurai}.

\subsubsection{Stochastic method}
based on the idea that it is better to start the execution at a low speed and accelerate the execution later when needed than to start with a high speed and reduce the speed later when slack time is found. By starting at a low speed, if the task finishes earlier than its WCET, it does not need to execute at a high speed. Theoretically, if the probability density function of execution times of a task is known \emph{a priori}, the optimal speed schedule can be computed \cite{gruian}.  Under the stochastic method, the clock speed is raised at specific time instances, regardless of the execution paths taken. The stochastic intra-task DVS may not utilize all the potential slack times.

\subsection{Inter-task DVS algorithms}
Inter-task DVS algorithms exploit the ``run-calculate-assign-run'' strategy to determine the supply voltage, which can be summarized as follows: (1) run a current task, (2) when the task is completed, calculate the maximum allowable execution time for the next task, (3) assign the supply voltage for the next task, and (4) run the next task.  Most inter-task DVS algorithms differ during step (2) in computing the maximum allowed time for the next task $\tau$ which is the sum of WCET of $\tau$ and the slack time available for $\tau$.

A generic inter-task DVS algorithm consists of two parts: slack estimation and slack distribution.  The goal of the slack estimation part is to identify as much slack times as possible while the goal of the slack distribution part is to distribute the resulting slack times so that the resulting speed schedule is as uniform as possible. Slack times generally come from two sources; \emph{static slack times} are the extra times available for the next task that can be identified statically, while {dynamic slack times} are caused from run-time variations of the task executions.

\subsubsection{Slack estimation methods}
\emph{Static slack estimation Maximum constant speed:} One of the most commonly used static slack estimation methods is to compute the maximum constant speed, which is defined as the lowest possible clock speed that guarantees the feasible schedule of a task set \cite{shin}.  For example, in EDF scheduling, if the worst case  processor utilization (WCPU) $U$ of a given task set is lower than 1.0 under the maximum speed $f_{max}$, the task set can be scheduled with a new maximum speed $f'_{max} = U \cdot f_{max}$. Although more complicated, the maximum constant speed can be statically calculated as well for RM scheduling \cite{shin, gruian}.

@@

\emph{Stretching to NTA:} even though a given task set  is scheduled with the maximum constant speed, since the actual execution times of tasks are usually much less than their WCETs, the tasks usually have dynamic slack times.  One simple method to estimate the dynamic slack time is to use the  arrival  time  of  the  next  task \cite{shin}. (The  arrival  time of the next task is denoted by NTA.) Assume that the current task $\tau$ is scheduled at time $t$. If NTA of $\tau$ is later than $(t + \textnormal{WCET}(\tau))$, task $\tau$ can be executed at a lower speed so that its execution completes exactly at the NTA.

\emph{Priority-based slack stealing}: this method exploits the basic  properties of  priority-driven scheduling such as  RM and EDF. The basic idea is that when a higher-priority task completes its execution earlier than its WCET, the following lower-priority tasks  can use the  slack time from the completed higher-priority task.  It is also possible for a higher-priority task to utilize the slack times from completed lower-priority tasks.  However, the latter type of slack stealing is computationally expensive to implement precisely.  Therefore, the existing algorithms are based on heuristics \cite{aydin, kim}.

\emph{Utilization updating:} the actual processor utilization during run time is usually lower than the worst case processor utilization. The utilization updating technique estimates the required processor performance at the current scheduling point by recalculating the expected worst case processor utilization using the actual execution times of completed task instances \cite{pillai}.   When the processor utilization is updated, the clock speed can be adjusted accordingly. The main merit of this method is its simple implementation, since only the processor utilization of completed task instances have to be updated at each scheduling point.

\paragraph{Slack distribution methods} In distributing slack times, most inter-task DVS algorithms have adopted a greedy approach, where all the slack times are given to the next activated task. This  approach is not an optimal solution, but the greedy approach is widely used because of its simplicity.

\section{Representative Algorithms} \label{hard-rt-algs}
Many DVS algorithms have been proposed or developed, especially for hard real-time systems. Since lowering the supply voltage also decreases the maximum achievable clock speed, various DVS algorithms for hard real-time systems have the goal of reducing supply voltage dynamically to the lowest possible level while satisfying the tasks' timing constraints\cite{kim-2}. The following is a representative list of DVS algorithms that have been proposed for addressing energy efficiency on mobile devices: two of which are based on the RM scheduling policy, while the other algorithms are based on the EDF scheduling policy, the two most widely used real-time system models\cite{w-s-liu}. Two algorithms were selected for intra-task DVS, one from path-based intra-task DVS algorithms, and the other from stochastic methods.

@@

Shin's intra-task DVS algorithm \cite{shin-kim-lee} and Gruian's algorithm \cite{gruian} are used as representative intra-task DVS algorithms which use path-based method and the stochastic method, respectively.

\section{Results} \label{sec:results}
In \cite{kim-2} a performance comparison of most of the algorithms presented in section \ref{hard-rt-algs} was made using a simulation environment named SimDVS. This work is relevant because it is argued that although each DVS algorithm is shown to be quite effective in reducing the energy/power consumption of a target system under its own experimental scenarios, these DVS algorithms have not been quantitatively evaluated under a unified framework, making it a difficult task for low-power embedded system developers to select an appropriate DVS algorithm for a given application/system.

\paragraph{Impact of the number of task on the energy consumption} as the number of tasks increases,  the  energy  efficiency  of \texttt{lppsEDF}, \texttt{lppsRM}, and \texttt{ccRM} that only use the Stretching-to-NTA technique do not significantly improve, while that of the other more aggressive inter-task DVS algorithms improves significantly when compared with the energy consumption of an applicacion running on a DVS-unaware system with a power-down mode only. In the Stretching-to-NTA method, the slack time that can be exploited is limited to the time between the completion of a task instance and the arrival time of the next task instance, which is largely independent of the number of tasks in the system. For the other inter-task DVS algorithms, since the slack times can be taken from any completed task instance, as the number of task increases, each task has more slack sources and can be scheduled with a lowered clock speed.

\paragraph{Worst case processor utilization of task set} except for \texttt{lppsEDF}, the energy consumption of inter-task DVS algorithms increases as a linear function of WCPU of a task  set. For \texttt{lppsEDF}, the energy consumption increases faster than a linear function of WCPU of a task set. This indirectly indicates that the dynamic slack estimation method of \texttt{lppsEDF} is not very effective. 
\texttt{lppsEDF} shows better energy efficiency than \texttt{ccEDF} when WCPU is less than 0.7. In \texttt{ccEDF}, the clock speed is determined using the actual processor utilization at the scheduling point.  Since the actual processor utilization increases when a low-speed task instance completes its execution, the next task instance needs to be executed in a higher speed. Such  voltage fluctuation occurs more often as the WCPU decreases. Thus, as the WCPU decreases, the energy efficiency of ccEDF becomes worse than that of \texttt{lppsEDF}. The results for \texttt{lppsRM} and \texttt{ccRM} are very similar to that of \texttt{lppsEDF}.

\paragraph{Speed bound} the energy efficiency of \texttt{AGR} and \texttt{lpSEH} is very close to the theoretical lower bound\footnote{Determined by the Yao's algorithm in \cite{yao}} when the speed bound factor is near 0.5. For the aggressive inter-task DVS algorithms, the energy efficiency is highest when the speed bound factor was set to average case processor utilization (ACPU). When the selected speed bound factor is close to ACPU $(= 0,55 \times \textnormal{WCPU})$, the best energy effciency is achieved for \texttt{laEDF}. For \texttt{ccEDF} this trend does not hold. For RM inter-task DVS algorithms, the performance gap between the energy efficiency and that of the theoretical lower bound was roughly 35$\sim$40\%.

@@

\subsection{Performance evaluation of Intra-Task DVS algorithms}
The two representative intra-task DVS algorithms perform quite differently depending on available slack times. When the relative energy consumption ratio of intraGruian over intraShin was measured, if the ratio is larger than 1, intraGruian performs better than intraShin. When  the  slack  ratio  is  less  than  1.2, intraShin outperforms intraGruian because intraShin spends more time in the lower speed region than intraGruian. When the slack ratio is increased, intraGruian spends more time in the lower speed region than intraShin.

\section{Related Work} \label{sec:rel-work}
Although this work is focused in hard real-time scheduling algorithms, CPU scheduling and energy/power efficiency of soft real-time systems and mobile devices has been addressed as well. In \cite{yuan-nahrstedt-1} a framework to integrate DVS into soft real-time scheduling for open mobile systems is presented. Its goal is to achieve energy saving DVS while preserinf resource guarantees of soft real-time scheduling. In \cite{yuan-nahrstedt} an enery efficient soft real-time CPU scheduler intended for mobile multimedia systems named \emph{GRACE-OS} is presented.

In \cite{kim-2} hybrid methods are discussed as well. By using this methods researchers try to answer quest of whether hybrid DVS algorithms will perform better than pure inter-task DVS or pure intra-task DVS. The study indicates that the performance of a HybridDVS algorithm can be better than a pure intra-task DVS algorithm or a pure inter-task DVS algorithm. However, the differences in energy efficiency depend on the characteristics of both the intra-task DVS and the inter-task DVS components used in the HybridDVS algorithm.

@@

\section{Conclusions} \label{sec:conclusion}
In this paper, real-time scheduling principles and algorithms and their impact on energy/power consumption were discussed. A high level introduction to the most popular hard real-time scheduling algorithms such as EDF and RM are given and later we also introduced variations of these real-time algorithms in the context of Dynamic Voltage Scaling, which allows devices to bring good performace while having improved battery consumption.

According to the results, the efficiency of the algorithms vary depending of the impact of conditions such as number of tasks or worst case processor utilization, but even under those conditions improvements are experienced which maken them suitable for scenarios when energy concerns are required.

@@

\bibitem{lee-sakurai}S.~Lee and T.~Sakurai. \emph{Run-time Voltage Hopping for Low-power Real-Time Systems}. In Proceedings of the 37th Design Automation Conference, pages 806–809, June 2000
